---
title: "Covid-19: Algeria"
description: "Policy Measures: Effective or ineffective"
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    theme: theme.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(rmarkdown)
```

```{r dataDownload}

covid_stats <- read_csv("data/owid-covid-data.csv")
measures <- readxl::read_excel("data/acaps_covid19_government_measures_dataset_0.xlsx",2)

```

```{r dataWrangling}

#Filter for Algeria
covid_stats_algeria <- covid_stats %>%
  filter(location == "Algeria") %>%
  select(date, new_cases, new_deaths, total_cases, total_deaths)

measures_algeria <- measures %>%
  filter(COUNTRY == "Algeria") %>%
  select(LOG_TYPE, CATEGORY, MEASURE, COMMENTS, DATE_IMPLEMENTED, SOURCE, SOURCE_TYPE, LINK)

covid_algeria <- full_join(covid_stats_algeria, measures_algeria, by = c("date" = "DATE_IMPLEMENTED"))
```

## Data Description

Data sets used in this report come from two main sources - 

1.    Collection of the COVID-19 data maintained by [Our World in Data](https://github.com/owid/covid-19-data/tree/master/public/data).
2.    [COVID-19 Government Measures Dataset](https://www.acaps.org/covid-19-government-measures-dataset) by ACAPS.

It contains a collection of data with `r length(covid_stats)` variables for COVID-19 dataset.
The data is available from `r covid_stats_algeria$date[1]` to `r covid_stats_algeria$date[550]` upon filtering for Algeria. The data sets are then filtered to select only the appropriate variables. Out of 62 variables in the COVID-19 dataset, only 5 were selected while 8 out of the 18 variables in the government measures dataset were selected. These tables are then merged together using the full join technique to get our final working dataset.

**Descriptions of variables -**
```{r}

k <- data.frame(Variables = names(covid_algeria),
                Description = c("Date values",
                                "New cases recorded on date",
                                "New deaths recorded on date",
                                "Total number of cases recorded till date from start of pandemic",
                                "Total number of deaths recorded till date from start of pandemic",
                                "Introduction / extension of measures, Phase-out of measures",
                                "Movement restrictions, Public- health measures, Social and economic measures, Lockdown, Humanitarian exemptions",
                                "Measures types and categories",
                                "Description of measure taken with further details",
                                "Source organization",
                                "Source organization type",
                                "Link to source website"))

paged_table(k, options = list(rows.print = 5, cols.print = 2))
  
```





## Citations

Hasell, J., Mathieu, E., Beltekian, D. et al. A cross-country database of COVID-19 testing. Sci Data 7, 345 (2020). https://doi.org/10.1038/s41597-020-00688-8

Covid-19 government measures dataset. ACAPS. (2020, December 10). https://www.acaps.org/covid-19-government-measures-dataset. 





