---
title: "Covid19: Austria"
description: |
  A Tale of Two Articles
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(lubridate)
library(mapdeck)
library(ggstream)
library(plotly)
```

## Introduction

Coronavirus disease (COVID-19) is an infectious disease caused by a newly discovered coronavirus. It began spreading in earnest in the early 2020's and has led to a global crisis of never-seen before proportions. People are being forced to stay indoors to reduce risk of exposure and spread of the disease.

Austria, a country in the Europe has also had to endure this crisis.

This blog post is aimed at investigating a few key articles that became breaking news during the time and checking the veracity of their claims.

## The First Article

```{r data1}

wave1 <- tibble(Country = c("Austria", "Iceland", "Denmark", "Norway", "Finland", "Germany"),
                Cases = c(1, 8, 4, 10, 6, 14),
                Dates = c("07/03/2020", "04/03/2020", "08/03/2020", "09/03/2020", "09/03/2020", "12/03/2020"),
                lat = c(47.516232, 64.963051, 56.263920, 60.472023, 61.924110, 51.1657),
                long = c(14.550072, -19.020836, 9.501785, 8.468946, 25.748152, 10.4515),
                lat_from = c(47.516232, 47.516232, 47.516232, 47.516232, 47.516232, 47.516232),
                long_from = c(14.550072, 14.550072, 14.550072, 14.550072, 14.550072, 14.550072),)


covid_stats <- read_csv("../../data/owid-covid-data.csv")
```



```{r map1}
key = "pk.eyJ1Ijoic21yaXRpYmhhdC1tb25hc2giLCJhIjoiY2t0anJjZW50MG41djJ3cW40cGpkMnp3eCJ9.trgg-yv2Femn9glwxksBVQ"

wave1 <- wave1 %>%
  mutate(cases_text = paste0("Number of cases: ", Cases))

mapdeck(token = key, style = mapdeck_style("light"), pitch = 85 ) %>%
  add_arc(
    data = wave1,
    layer_id = "arc_layer",
    origin = c("long_from", "lat_from"),
    destination = c("long", "lat"),
    stroke_from = "#FFB344",
    stroke_to = "#00A19D",
    stroke_width = 10,
    tooltip = "cases_text",
    auto_highlight = TRUE
  ) %>%
  add_title(title = "Spread of Covid-19 from Ground Zero", layer_id = "title")
```

```{r viz2}

covid_stats_eu <- covid_stats %>%
  filter(location %in% wave1$Country) %>%
  select(date, new_cases, new_deaths, location) %>%
  filter(date <= "2020-04-05")

cols <- c("#FFB400", "#FFC740", "#C20008", "#FF020D", "#13AFEF", "#4276ac")

p <- ggplot(covid_stats_eu, aes(date, new_cases, fill = location, label = new_cases)) +
  geom_stream(color = 1, lwd = 0.25, bw = 1) +
  scale_fill_manual(values = cols) +
  theme_minimal()


ggplotly(p)

```



## The Second Article


```{r}
covid_stats_au_1 <- covid_stats %>%
  filter(location == "Austria") %>%
  select(date, new_cases, new_deaths) %>%
  filter(date >= "2020-11-06" & date <= "2020-12-06")

covid_stats_au_2 <- covid_stats %>%
  filter(location == "Austria") %>%
  select(date, new_cases, new_deaths) %>%
  filter(date >= "2020-12-06")
```

```{r}

```




## Citations


Hersche, D. I. B., Kern, U.-P. D. W. V., Stuber- Berries, D. N., Rohrer, D. R., Trkola, D. A., &amp; Weber, U.-P. D. K. (2020, October 12). BERICHT DER UNABHÄNGIGEN EXPERTENKOMMISSION MANAGEMENT COVID-19-PANDEMIE TIROL. Vienna; Consumer Protection Association (VSV). https://www.verbraucherschutzverein.eu/wp-content/uploads/2021/08/Expertenkommission-Bericht.pdf

Guardian News and Media. (2020, November 27). 'A Christmas not like others': Europe wrestles with FESTIVE Covid rules. The Guardian. Retrieved September 15, 2021, from https://www.theguardian.com/world/2020/nov/27/christmas-not-like-others-europe-wrestles-festive-covid-rules. 

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686

Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with lubridate. Journal of Statistical Software, 40(3), 1-25. URL https://www.jstatsoft.org/v40/i03/.

Hao Zhu (2021). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax. R package version 1.3.4. https://CRAN.R-project.org/package=kableExtra

Yihui Xie (2021). knitr: A General-Purpose Package for Dynamic Report Generation in R. R package version 1.31.

Yihui Xie (2015) Dynamic Documents with R and knitr. 2nd edition. Chapman and Hall/CRC. ISBN 978-1498716963

Yihui Xie (2014) knitr: A Comprehensive Tool for Reproducible Research in R. In Victoria Stodden, Friedrich Leisch and Roger D. Peng, editors, Implementing Reproducible Computational Research. Chapman and Hall/CRC. ISBN 978-1466561595








